Empty Homes V1

### 2017-03-20 EmptyHomesV1.Rmd
I haven't done an update in a while so all kinds of changes have been made. A lot of the changes are to do with making a map at ward level not lsoa level due to problems with getting data from the councils.

```{r}
packages <- c("tidyverse", "spdep", "igraph", "maptools")

#you could replace a lot of the packages with just tidyverse but some people have problems installing it.
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(packages, library, character.only = TRUE)

#Set up file system to read the correct folders this switches between aws and windows mode
basewd <-"~/Dropbox/SSE/Empty Homes"

DataFolder <- file.path(basewd, "Data")
ShapeFiles <- file.path(basewd,"ShapeFiles/statistical-gis-boundaries-london/ESRI")
Figures <- file.path(basewd, "Figures")

select <-dplyr::select
```

@@

>>>>

It would have been sensible to separate council and private housing

```{r setup}
setwd( "~/Dropbox/SSE/Empty Homes/EmptyHomesCode/SubCode")
source("Setup.R")
```

#Load completed data

```{r}
setwd("/home/jonno/Dropbox/SSE/Empty Homes/Regions-Complete/Region London")
list.files()

test <- read.xlsx("SuttonExemptionsLSOA.xlsx")
ken1 <- read.xlsx("Copy of Kensington and ChelseaCTBandsLSOA.xlsx" )
ken2 <- read.xlsx("Copy of Kensington and ChelseaExemptionsLSOA (002).xlsx")
```

@@

Council tax bands from VOA
https://www.gov.uk/government/statistics/council-tax-stock-of-properties-2016

```{r PopandHomes}
setwd(DataFolder)

CTtaxstock<- read_csv("Table_CTSOP1.1_2016.csv")

pop <- read_csv("SAPE2.csv")

EW <- inner_join(select(CTtaxstock, ECODE, ALL_PROPERTIES),select(pop, `Area Codes`, `All Ages`),
          by=c("ECODE" = "Area Codes")) %>% setNames(c("ECODE", "Homes", "Pop")) 
```

#Padstow example
```{r}
test2 <- CTtaxstock 
pop2 <- pop %>% filter(grepl("Cornwall 008",X3))

#padstow
pad <- left_join(select(test2, ECODE, ALL_PROPERTIES), select(pop2, `Area Codes`, `All Ages`),
          by=c("ECODE" = "Area Codes")) %>% 
  mutate(diff = ALL_PROPERTIES-`All Ages`, ratio = ALL_PROPERTIES/`All Ages`) %>% filter(complete.cases(.))

#Load the LSOA shape data for EW

```{r}
setwd(LSOAshapedata)

shape <- readOGR(dsn = list.files(pattern = "shp"))

# Load Post code data

This chunk checks the exiatane of the postcode to lsoa lookup if it doesn't find it, the second part of the code is run that creates the lookup. Creating the lookup takes quite a lot of time so it is good to only have to do it once.
```{r}

if(file.exists(file.path(DataFolder,"PstCdLSOA.rds"))){
 PstCdLSOA.raw <- readRDS(file.path(DataFolder,"PstCdLSOA.rds"))
} else {
  source(file.path(CommonCode, "PostcodeLookup.R"))
}

Bind the ward names and LAD names to the lsoa data
```{r}
setwd(PostcodeLookups)
lsoa <- read_csv("PCD11_OA11_LSOA11_MSOA11_LAD11_EW_LU_aligned_v2.csv") %>%
  select(LSOA11CD:LAD11NMW) %>% distinct(., LSOA11CD, .keep_all = TRUE)

setwd(AddGeog)
wardnames <- read_csv("Ward_to_Local_Authority_District_to_County_to_Region_to_Country_December_2016_Lookup_in_United_Kingdom_V2.csv") %>% setNames(c("Admin_ward_code", names(.)[-1]))

#uses the dataframe from the CreateExceltemplates to get the ward names and the lsoa together
wardlsoa <- PstCdLSOA.raw %>% distinct(lsoa11cd, .keep_all = TRUE) %>% 
  select(lsoa11cd, Admin_ward_code)

EW2 <- inner_join(EW, lsoa, by = c("ECODE" = "LSOA11CD") ) %>% 
      left_join(., wardlsoa, by = c("ECODE"="lsoa11cd") ) %>% 
  left_join(., select(wardnames, Admin_ward_code, WD16NM), by = "Admin_ward_code") %>%
  group_by(Admin_ward_code) %>%
  mutate(WardHomes = sum(Homes),  WardPop = sum(Pop)) %>% ungroup
rm(wardlsoa)
rm(wardnames)
```

Convert shape file to dataframe and join in the LSOA data
```{r}
setwd(LSOAshapedata)

shape <- readOGR(dsn = list.files(pattern = "shp"))

AG <- fortify(shape, region = "objectid")

AG <- AG %>% mutate(id = as.integer(id)) %>% 
  left_join(., shape@data, by = c("id" = "objectid")) %>% 
  left_join(., EW2, by = c("lsoa11cd" = "ECODE"))

#I should probably join the homes and pop at the same time I join the council tax data

other borough
```{r}
setwd("/home/jonno/Dropbox/SSE/Empty Homes/Regions-Complete/Region London")
list.files()

test <- read_excel("Copy of Haringey_Exemptions_Discounts_LSOA.xlsx")[,c(1,4:6)] %>% setNames(make.names(names(.)))

test2 <- test %>% group_by(LSOA_CODE, Exemption.type) %>% summarise(Counts = n())

test2b <- test %>% 
  mutate(type2 = ifelse(grepl("Empty", Exemption.type), "Empty", Exemption.type)) %>%
  group_by(type2) %>% summarise(Counts = n())

test2b <- test %>% 
  mutate(type2 = ifelse(grepl("Empty", Exemption.type), "Empty", Exemption.type)) %>%
  group_by(type2, LSOA_CODE) %>% "Hammersmith.xls"  
  summarise(Counts = n()) %>% 
  spread(., type2, Counts, fill = 0)

test3 <- test2b %>% left_join(., select(EW2, ECODE:Pop), by = c("LSOA_CODE" = "ECODE")) %>%
   mutate(EmptPerc = Empty/Homes)

Load lambeth data
```{r}
setwd("/home/jonno/Dropbox/SSE/Empty Homes/Regions-Complete/Region London")
list.files()

test <- read_excel("LambethExemptionsLSOA.xlsx")[,1:4] %>% setNames(make.names(names(.)))

test2b <- test %>% 
  mutate(type2 = ifelse(grepl("Empty", Exemption.type), "Empty", Exemption.type)) %>%
  group_by(type2, LSOA_CODE) %>% 
  summarise(Counts = n()) %>% 
  spread(., type2, Counts, fill = 0)

test3 <- test2b %>% left_join(., select(EW2, ECODE:Pop), by = c("LSOA_CODE" = "ECODE")) %>%
   mutate(EmptPerc = Empty/Homes)

Filter to lambeth and plot
```{r}
AG2 <- AG %>% filter(LAD11NM == "Lambeth") %>% 
  left_join(., test2b, by = c("lsoa11cd" = "LSOA_CODE")) %>% mutate(EmptPerc = Empty/Homes)

ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group, fill = EmptPerc), 
colour = alpha("darkred", 1/2), size = 0.7) + scale_colour_gradient2() +
  ggtitle("Lambeth coloured by areas of highest % empty Homes")

Filter to RBKC
```{r}

AG2 <- AG %>% filter(LAD11NM == "Kensington and Chelsea")

#manually map ward codes to empty and missing homes

WardDat <- data.frame(Admin_ward_code = unique(AG2$Admin_ward_code), 
           Wardname = c("Abingdon","Brompton and Hans Town", "Courtfield", "Stanley", "Campden", "Colville", "Earl's Court", 
                      "Redcliffe", "Chelsea Riverside", "Kensal Green", "Golborne", "Royal Hospital", "Holland", "Notting Dale",
                      "Norland", "St. Helen's", "Pembridge", "Queen's Gate", "Dalgarno"),
           LowUse = c(744,1932,1060,593,664,330,572,587,371,0, 196, 829, 640,106,237, 149,331,846,87))

Chelsea <- EW2 %>% filter(LAD11NM == "Kensington and Chelsea") %>% 
  select(Admin_ward_code:WardPop) %>%
  distinct(Admin_ward_code, .keep_all=TRUE) %>% 
  left_join(., WardDat) %>% ungroup %>% mutate(PercEmpt = LowUse/WardHomes)

AG2 <- AG2 %>% left_join(., WardDat, by = "Admin_ward_code") %>% 
  filter(Wardname != "Kensal Green")

ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group, fill = vallow), 
colour = alpha("darkred", 1/2), size = 0.7) + scale_colour_gradient2() 
```

#get prices

```{r}

setwd(DataFolder)

test <-read_csv("pp-2016.csv", col_names = FALSE )

#quick and dirty loses half the sales due to missing the postcode
test <- test %>% left_join(., PstCdLSOA.raw, by = c("X4"="Postcode")) %>% filter(!is.na(lsoa11cd))

test2 <- test %>% group_by(Admin_ward_code) %>% summarise( MeanPrice = mean(X2), counts =n() )

WardDat2 <- WardDat %>% left_join(., test2) %>% mutate(vallow = LowUse*MeanPrice) %>%
  filter(!is.na(MeanPrice))

sum(WardDat2$vallow)/sum(WardDat2$LowUse)

Lest <- test %>% filter(X14 =="LEICESTERSHIRE")

Lest %>% filter(X2<15000000) %>% ggplot(., aes(x= X2)) + geom_density()

%>% group_by(X13) %>% summarise(mean = mean(X2))

Melt <- test %>% filter(X13 == "MELTON")

test %>% filter(grepl("chelsea", X13, ignore.case = TRUE)) %>%
  filter(X2<3000000) %>% ggplot(., aes(x= X2)) + geom_density()

# Using the semi data

```{r}
#Lambeth

setwd("/home/jonno/Dropbox/SSE/Empty Homes/Regions-Complete/Region London")

Lambeth <- read_excel("LambethExemptionsLSOA.xlsx")[,1:4] %>% setNames(make.names(names(.))) %>% 
  mutate(type2 = ifelse(grepl("Empty", Exemption.type), "Empty", Exemption.type)) %>%
  group_by(type2, LSOA_CODE) %>% 
  summarise(Counts = n()) %>% 
  spread(., type2, Counts, fill = 0) %>% 
   left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% 
  group_by(Admin_ward_code) %>% mutate(Empty = sum(Empty), LowUse = Empty) %>% ungroup

LAMDAT <- AG %>% filter(LAD11NM == "Lambeth") %>% 
  left_join(., select(Lambeth, LSOA_CODE, LowUse, Empty), by = c("lsoa11cd" = "LSOA_CODE"))

#Sutton

Sutton <- read_excel("SuttonExemptionsLSOA.xlsx"  )[,1:6]

#ward folder
setwd(SemiLondon)
list.files()

wardnames <- EW2 %>% group_by(Admin_ward_code, WD16NM) %>% summarise(counts = n(), LAD11NM = first(LAD11NM), LAD11CD = first(LAD11CD)) %>% arrange(WD16NM)

test <- wardnames %>% filter(grepl("Kensington", LAD11NM))

test <- EW2 %>% filter(LAD11NM == "Wandsworth")

#Chelsea
Chelwards <- wardnames %>% filter(grepl("Kensington", LAD11NM))

Chelsea <- read_excel("FOI2017-0143.xlsx" ) %>% setNames(make.names(names(.))) %>% slice(1:18) %>%
  mutate(Admin_ward_code = Chelwards$Admin_ward_code[c(1:10,12:17,19,18)]) %>% group_by(X.Ward) %>%
  mutate(Empty = sum(Empty..unfurnished, Empty..works, Job.related., Empty..50..premium., Moorings , na.rm = TRUE),
         LowUse = Empty + Second.Homes) %>% ungroup

Chelsea[is.na(Chelsea)] <-0
rm(Chelwards)

CHELDAT <-  AG %>% filter(grepl("Kensington", LAD11NM))%>% left_join(., select(Chelsea, Admin_ward_code, LowUse, Empty)) %>% 
  filter(WD16NM!= "Kensal Green")

#Hackney has provided Census Wards I'm not sure how useable it is. 
hackney <- read_csv("hackney.csv") %>% setNames(make.names(names(.))) %>% group_by(Exemption.Discount, Ward.Name) %>% summarise(counts = n()) %>% ungroup %>% spread(Exemption.Discount, counts, fill = 0)

#Hammersmith 
Hammersmith <- read_excel("Hammersmith.xls", skip = 1 ) %>% filter(!is.na(Ward) ) %>% 
  bind_cols(., filter(wardnames,LAD11NM == "Hammersmith and Fulham")) %>% select(-Ward) %>%
  mutate(LowUse = `Second (B)` + `Empty Unocc`+ LTE + Premium, Empty = LowUse - `Second (B)`)

Hammersmith2 <- EW2 %>% filter(LAD11NM == "Hammersmith and Fulham") %>% left_join(., select(Hammersmith, Admin_ward_code, LowUse, Empty))

HAMDAT <- AG %>% filter(LAD11NM == "Hammersmith and Fulham")%>% left_join(., select(Hammersmith, Admin_ward_code, LowUse, Empty))

#Wandsworth

wardnames %>% filter(LAD11NM == "Wandsworth")

Wandsworth <- read_excel("wandsworth2.xlsx",  skip = 3 ) %>% select(1,3) %>% 
  group_by(`Ptc Wardname`) %>% summarise(Empty = sum(`Count Current Occupancy Discounts`)) %>% ungroup %>% 
  arrange(`Ptc Wardname`) %>% 
  filter(!is.na(`Ptc Wardname`)) %>% 
  bind_cols(., wardnames %>% filter(LAD11NM == "Wandsworth") %>% 
              select(Admin_ward_code)) %>%
  mutate(LowUse = Empty)

Wandsworth2 <- EW2 %>% filter(LAD11NM == "Wandsworth") %>% left_join(., select(Wandsworth, Admin_ward_code, Empty))
WANDAT <- AG %>% filter(LAD11NM == "Wandsworth")%>% left_join(., select(Wandsworth, Admin_ward_code, LowUse, Empty))

#Merton
wardnames %>% filter(LAD11NM == "Merton")
Merton <- read_excel("merton1.xlsx" )[,-2] %>%setNames(make.names(names(.))) %>% slice(-1) %>% arrange(X) %>%
  filter(X != "Norbury", X!="Nightingale", X !="Stonecot") %>% 
    bind_cols(., wardnames %>% filter(LAD11NM == "Merton") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty =   Disc.Class.C.Vacant.dwelling + Disc.class.C.Long.term.empty + Premium.Levy, LowUse =  Disc.Class.A.Second.home + Disc.Class.B.second.home )

MERDAT <- AG %>% filter(LAD11NM == "Merton")%>% left_join(., select(Merton, Admin_ward_code, LowUse, Empty))

#Southwark

test <- wardnames %>% filter(LAD11NM == "Southwark")

Southwark <- read_excel("southwark.xlsx" , skip = 1) %>% setNames(make.names(names(.))) %>% 
  filter(Ward.codes != "TOTALS") %>% select(-Ward.codes) %>%
  mutate(WARDS= na.locf(WARDS)) %>% group_by(WARDS) %>%
  summarise_all(sum, na.rm = TRUE) %>% ungroup %>%
  arrange(WARDS) %>%
    bind_cols(., wardnames %>% filter(LAD11NM == "Southwark") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty =   Unoccupied.and.furnished.lets....0. + Job.related.dwelling...50..discount + Second.Home...unoccupied.and.furnished...0.+Empty.and.unfurnished...0. + Levy.on.long.term.empty...50..Premium + Uninhabitable.Works.required....0., LowUse =  Empty )

Wandsworth2 <- EW2 %>% filter(LAD11NM == "Southwark") %>% left_join(., select(Southwark, Admin_ward_code, Empty))

SOUDAT <- AG %>% filter(LAD11NM == "Southwark")%>% left_join(., select(Southwark, Admin_ward_code, LowUse, Empty))

#Harrow

test <- read_excel("HarrowCTBandsLSOA. 3118795.xlsx"  )[1:4]%>% setNames(make.names(names(.)))

#Kingston
test <- wardnames %>% filter(grepl("Kingston upon Thames", LAD11NM))

Kingston <- read_excel("Kingston1.xlsx"  )%>% setNames(make.names(names(.))) %>%
  arrange(X) %>% slice(1:16) %>%
  bind_cols(., wardnames %>% filter(LAD11NM == "Kingston upon Thames") %>% 
              select(Admin_ward_code)) %>%
  mutate(Empty = Empty...Furnished..0.. + Empty.Homes.Premium + LONG.TERM.EMPTY..0.. + Structural.Repair..0.., LowUse = Empty + X2nd.Home..0..)

KINDAT <- AG %>% filter(LAD11NM == "Kingston upon Thames")%>% left_join(., select(Kingston, Admin_ward_code, LowUse, Empty))

#TowerHamlets
test <- wardnames %>% filter(grepl("Tower", LAD11NM))

Tower <- read_excel("Towerhamlets.xlsx")[,c(1,3)] %>% 
  filter(!is.na(DiscClass)) %>%
  group_by(Ward, DiscClass) %>%
  summarise(count = n()) %>% ungroup %>%
  spread(., key =DiscClass, value = count, fill = 0 ) %>% 
  mutate(Empty = rowSums(.[,4:8]), LowUse = B+BE+Empty) %>% 
  slice(c(1:13, 15:17, 14, 18:20)) %>%
   bind_cols(., wardnames %>% filter(grepl("Tower", LAD11NM)))

TOWDAT <- AG %>% filter(grepl("Tower", LAD11NM)) %>% left_join(., select(Tower, Admin_ward_code, LowUse, Empty))

#Greenwhich
list.files()
wardnames %>% filter(grepl("Greenwich", LAD11NM))

Greenwich <- read_excel( "FOI-3348 ctfoi35report.xlsx" ) %>%  
  setNames(make.names(names(.))) %>%
  filter(grepl("EMPTY",Exempt.Discount.Description)) %>% 
  select(-Exempt.Code) %>%
  spread(key= Exempt.Discount.Description, value = Total, fill = 0) %>%
  mutate(Empty = rowSums(.[,-1]), LowUse = Empty) %>%
   bind_cols(., wardnames %>% filter(grepl("Greenwich", LAD11NM)))

GREDAT <- AG %>% filter(grepl("Greenwich", LAD11NM)) %>% left_join(., select(Greenwich, Admin_ward_code, LowUse, Empty))

Plot Map
```{r}
AG2 <- bind_rows(CHELDAT,HAMDAT, WANDAT, LAMDAT, MERDAT, 
                 SOUDAT, KINDAT, TOWDAT, GREDAT) %>% 
  mutate(LowUsePerc = LowUse/WardHomes, 
         EmptyPerc = Empty/WardHomes,
         EmptyPercRank = percent_rank(Empty/WardHomes))

ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group, fill = EmptyPerc), 
colour = alpha("darkred", 1/2), size = 0.1) + scale_colour_gradient2() +
  labs(title = "All Boroughs Percentage Empty") +
  theme(plot.title = element_text(hjust = 0.5))

AG2 %>% 
  ggplot(.)+ geom_polygon( aes(long, lat, group = group, fill =sqrt(EmptyPerc)), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2()

ggplot()+ geom_polygon(data=AG2, aes(long, lat, group = group, fill = LowUse/WardHomes), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2()

#single borough
GREDAT %>%
  mutate(LowUsePerc = LowUse/WardHomes, 
         EmptyPerc = Empty/WardHomes,
         EmptyPercRank = percent_rank(Empty/WardHomes)) %>%
ggplot(.)+ geom_polygon( aes(long, lat, group = group, fill = EmptyPerc), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2()

#Proper data Barrow

```{r}
setwd(file.path(RegionsComp, "Region North West"))

list.files()

test <- read_excel("Barrow-in-FurnessExemptionsLSOA.xlsx", sheet=1) %>% setNames(make.names(names(.))) %>% .[,1:4] %>% filter(Current.Discount.Type.Code !="SINGLE") %>% group_by(LSOA_CODE) %>%
  summarise(Empty = n())

Barrow2 <- test %>% left_join(., EW2, by = c("LSOA_CODE" = "ECODE")) %>% mutate(EmptyPerc = Empty/Homes)

test2 <- test %>% left_join(., AG, by = c("LSOA_CODE" = "lsoa11cd")) %>% mutate(EmptyPerc = Empty/Homes)

ggplot()+ geom_polygon(data=test2, aes(long, lat, group = group, fill = EmptyPerc), 
colour = alpha("darkred", 1/2), size = 0.2) + scale_colour_gradient2() + ggtitle("Barrow in Furness Emptyness %")
```
